Mixed language processing increases cross-language phonetic transfer in Bengali-English bilinguals

### 2021-10-19 draft_BLC/BLC_AM-ID_manuscript.tex
BLC reviewer comments

EFL University

Hyderabad\\

@@

School of languages and Linguistics

Jadavpur University

Kolkata\\

@@

\newpage
\begin{myparindent}{0pt}
Abstract:\\
This study investigates the phonetic outcome of mixed-language speech in Bengali and Indian English, towards understanding cross-language interaction in highly proficient bilinguals.  We compare spectral properties of L2 vowels \nt{\ae} (common to L1, L2) and \nt{2} (absent in L1) in unilingual vs mixed productions. Results reveal dynamic and asymmetrical changes in both vowels;  \nt{2} shifting towards a related L1 category \nt{a:}. We interpret this as a temporary increase in cross-language phonetic interaction during mixed-language use, leading to a shift towards L1 norms, evidencing transfer effects on L2 vowels for the first time. We elicit mixed productions through two tasks: cued picture-naming and code-switching, to assess if experimental paradigm independently influences the behavior under study. Results reveal no difference across paradigms. We discuss these findings in light of recent proposals about asymmetries in short-term phonetic interaction between sound systems and the postulated role of connected speech in introducing discursive factors to transfer studies.\\

@@

The rest of this section is structured as follows: \ref{bengali_english_in_india} provides background information on Bengali and Indian English focusing on vowel systems. \ref{causes} discusses two potential sources of transfer during language processing-- global co-activation of multiple languages vs the local act of switching between languages. Given the highly multilingual setting that characterizes this population, we note that the former is expected to be a constant feature and therefore any observed phonetic interaction is better understood as resulting from the latter. The design of the present study is motivated on the basis of this discussion. A logical consequence of this model is that such phonetic interaction must be highly localized. \ref{duration} elaborates on this by reviewing literature on the duration and `reversiblity' of transfer effects. A recurring pattern that emerges from existing research on short-term phonetic transfer is that both languages of the bilingual speaker are not equally affected, although the nature of these reported differences varies greatly across studies. In \ref{asymmetries}, we review some of these findings, focusing on proposed sources of asymmetries, in order to highlight the factors that mediate such phonetic interaction (or fail to do so). We argue that these results point to an urgent need for expanding the range of language pairs and sound categories examined in order to make meaningful generalizations about the nature of short-term phonetic transfer. Following from this, \ref{asymmetry between sounds} discusses reported asymmetries between different sound categories of a single language, and develops the hypotheses of the present study on the basis of these. In \ref{paradigms}, we motivate a secondary aim of this study: to verify the postulated differences between two paradigms for eliciting short-term phonetic interaction in an experimental setup, namely cued picture-naming and code-switching. \ref{questions_and_hypotheses} summarizes the research questions and hypotheses.

@@

Existing research has distinguished changes in category \textsc{representations} due to the acquisition of multiple sound systems (cf. Speech Learning Model \citep{flege1995second,flege2007language}; Perceptual Assimilation Model-L2 \citep{best2007nonnative}), from interaction during accessing, processing, or articulation of these categories (cf. transfer vs interference \citep{grosjean2012attempt}; competence vs performance interference \citep{paradis1993linguistic}). 
Given their transient nature, dynamic changes in production during mixed-language use are generally attributed to the latter, e.g. online processing costs \citep{olson2013bilingual,tsui2019impact,vsimavckova2015immediate}, language mode \citep{simonet2014phonetic}, context-awareness \citep{khattab2013phonetic}. What triggers this interaction? \cite{olson2016role} argues that while cross-language phonetic effects are largely measured at the point of language switch, it could have two potential sources:
\begin{enumerate}
	\item The local point of switch itself
	\item Global co-activation of two languages -- \textsc{bilingual language mode} \citep{grosjean1998studying} 
\end{enumerate}

@@

\subsubsection*{L1 vs L2 status} 
Flege's Speech Learning Model (SLM) \citeyearpar{flege1995second,flege2007language} posits that the sound categories of a bilingual speaker exist in a common phonological space, and therefore in principle, both L1 and L2 categories can influence one another. Thus, the patterns of cross-language transfer depend on how L2 phonemes are mapped in relation to the existing L1 categories. In non-switched production, this is evidenced through a pervasive L1 influence on non-native contrasts that are perceptually linked to an existing L1 contrast (cf. \textsc{equivalence classification}-- \cite{flege1984limits,flege1987production}), and the observation that both L1 and L2 sound systems of bilinguals differ from those of corresponding monolingual speakers \citep[e.g.][]{guion2003vowel}.

@@

The precise contribution of various (linguistic and extra-linguistic) factors and their interactions to the phonetics mixed-language use merits detailed investigation in its own right. In the present study, however, our focus is limited to verifying whether, in an experimental setting, the postulated discursive elements of a CS paradigm alter the extent of observable transfer compared to a paradigm that arguably lacks these--- cued picture-naming. We are less concerned with the exact nature of differences than with their comparability. While researchers have successfully employed each of these paradigms to induce phonetic interaction, comparison across studies is difficult since they also differ in other factors. We provide a starting point for such comparison by eliciting productions of the same tokens from the same set of participants, in both paradigms. A difference in results will suggest that the planning involved in connected speech affects transfer above and beyond the phonetic outcome of switching itself, which has implications that future work should consider.

\subsection{Research questions and hypotheses}\label{questions_and_hypotheses}
\begin{itemize}
	\item \textbf{L2 vowels}: Are L2 vowel categories of proficient bilingual speakers affected by dynamic phonetic transfer during mixed-language production, as evidenced by differences in the spectral properties of L2 vowels in switched and non-switched tokens?
	\item \textbf{Asymmetry}: If so, we expect an asymmetry between the vowels in the extent and direction of shift. Based on comparisons with Bengali vowels from \cite{dutta2021} we expect \nt{2} to be lowered and fronted in the mixed condition. \nt{\ae} is expected to show lowering, but to a lesser extent than \nt{2}, as it is in a part of the IE vowel space that has greater vowel density.
	\item \textbf{Paradigm}: Is there an observable difference between the extent of transfer in a cued picture-naming vs a code-switching paradigm? 
\end{itemize}

@@

10 Bengali-English bilingual speakers (male=5, age range 19 to 28), completed the tasks. All the participants speak other languages in addition to Bengali and English. At the time of recording, they were living in or around the EFL University campus in Hyderabad, which has a linguistically diverse population. Speakers were selected on the basis of responses to a Language Background Questionnaire to ensure comparable LSRW skills in L1 and L2 across participants, and were compensated for their time.

\subsection{Stimuli}
Twenty monosyllabic English words containing the vowels \nt{2} or \nt{ae} were selected as target words. To minimize coarticulatory effects on the vowel, the pre-vocalic consonants were all voiced plosives (\nt{b} or \nt{d}). 10 monosyllabic English words that do not contain the target vowel served as filler items. We avoided English words that are lexicalized loanwords in Bengali. %The list of words can be found in sec.\ref{appendix}.

@@

\begin{enumerate}[]
	\item Cued picture-naming: Participants were presented with slides in the following sequence:
	\begin{itemize}
		\item Language cue- a word in either English or Bengali orthography
		\item A picture displayed for 50ms; named by the participant in either English or Bengali, depending on the language cue 
		\item Target -- an English word; read out by the participant
		\item Distracter math problem
	\end{itemize}
	Since all the target words were in English, the language cue was used to manipulate the \textit{preceding} word, giving target words in non-switched (unilingual) and switched (mixed) utterances. The orthographies of English and Bengali are visually distinct, and were therefore used to cue language.

@@

\end{enumerate}
\subsection{Acoustic analysis}\label{analysis}
The first iteration of each task was treated as a trial, and excluded from further analyses. An additional 67 tokens (2.79\%) were excluded due to errors in presentation/production, giving 2333 target word tokens for analysis. Recordings were manually segmented and annotated in Praat \citep{boersma2016praat}, and the first and second formant frequencies (F1 and F2) of the target vowels \nt{2, ae} were measured at five points-- in 10\% increments starting at 5\% into the vowel. This gave measures of F1 and F2 at 5\%, 15\%, 25\%, 35\% and 45\% into the vowel, allowing us to observe any dynamic changes in the extent of transfer over the course of producing the vowel.

@@

Vowel quality and gender are expected to systematically affect F1. Thus, we began with a model that had Vowel and Gender as fixed effects, random intercepts for Subject and Item, and random slopes for each within-unit variable that was relevant to our hypotheses-- by-subject slopes for Task, Context, and Vowel; by-item slopes for Task and Context. Next, adding Context as a fixed effect significantly improved the model fit ($\chi^2(1) = 5.54, p= .01854$), suggesting that the experimental condition affects vowel height. Following from our hypothesis \textbf{Asymmetry}, we expected the extent of change in vowel height between the conditions to be mediated by the identity of the vowel. Thus, we added a two-way interaction factor between Context and Vowel (Context*Vowel). As expected, this was found to significantly improve model fit ($\chi^2(1) = 12.61, p= .0003837$). To assess whether the experimental paradigm affected F1, we added Task as a fixed effect. This marginally improved the model ($\chi^2(1) = 3.74, p= .05302$). Towards addressing our research question \textbf{Paradigm}, we wanted to assess whether the extent of shift differs across the two paradigms. However, no significant improvement was achieved by adding an interaction effect between Task and Context. Subsequent additions of other interactions, and removing each of the fixed effects in turn, did not lead to any further improvement.

@@

With regard to our first research question \textbf{L2 vowels}, the results confirm that the L2 vowels of proficient bilinguals can be affected by mixed-language processing. Both vowels in this study showed systematic shifts in the mixed-language condition compared to baseline unilingual productions. In the context of an ICM--based analysis of between-language asymmetries in phonetic transfer as proposed by \cite{olson2013bilingual}, this is a significant finding: it demonstrates that the postulated lower switch cost on an L2 does not entirely mask transfer effects on L2 vowels. Moreover, our findings differ from \cite{tsui2019impact} who report on the basis of VOT measurements that balanced bilinguals do not evidence an effect of language-switching. \citeauthor{tsui2019impact} attribute the absence of transfer effects to greater experience with language switching in these participants, leading to lower switch-costs in both languages. We find that even proficient bilinguals with extensive experience of language switching demonstrate cross-language transfer effects in vowel production. Thus, following from the discussion in sec. \ref{asymmetries}, it might be fruitful for future studies to focus on vowels as a potential site for transfer. This would both expand the empirical scope of the existing research (cf. sec. \ref{asymmetry between sounds}) and avoid confounding factors arising from independent differences between long- and short-lag VOT languages.

@@

Our third research question \textbf{Paradigm} concerned the effect of experimental paradigm on the phonetic outcome of mixed-language use. \cite{olson2013bilingual} proposes that in both scripted and spontaneous code-switching paradigms, ``discursive properties may partially drive the production of code-switches, masking underlying effects of interaction", and thus proposes cued picture-naming as a preferable alternative. If such discursive properties independently affect the phonetic outcome of language mixing in an experimental setup, this calls to question the comparability of results across studies using these different paradigms. We verified this by comparing productions of the same items from the same group of participants in a scripted CS paradigm and a cued picture-naming paradigm. We found no significant interaction of Task either with Context, or with the Vowel*Context interaction, suggesting that the extent of phonetic transfer was not affected by the paradigm. 
There are at least two plausible reasons for this-- (i) it is possible that the planning effects of CS do not translate to an increase/decrease in observable transfer at all. This would mean that the pre-switch effects observed, for example, in \cite{muldner2019phonetics} and \cite{bullock2009trying} are better interpreted as resulting from some other aspect of the experimental setup, than the act of code-switching itself. While planning could additionally affect the words preceding the switch, the effects on the switched token itself do not change. (ii) Our study used a frame sentence to induce code switching, and not all the items led to pragmatically meaningful sentences. Thus, another possible explanation is that the discursive effects discussed by \cite{olson2013bilingual} are triggered by more naturalistic paradigms where attention to meaning and interlocutor are important. More within-study comparisons with less structured paradigms, and spontaneous code-switching data, can clarify this. The results of the present study indicate that for the purposes of observing dynamic phonetic transfer in an experimental setup, elicitation through cued-picture-naming vs (controlled) code-switching does not significantly alter the behavior under study. Thus, the choice of paradigm can be based on other considerations. Moreover, this suggests that (all else being equal) the results from such studies can be compared directly.

@@

\section{Conclusion}
This study examined the phonetic outcome of mixed-language production in a group of highly proficient bilingual speakers of Bengali and Indian English. Specifically, we examined the spectral quality of two L2 vowel categories \nt{2} and \nt{\ae} produced in two experimental conditions: unilingual (English) and mixed (switching from Bengali to English).
Comparisons of F1 and F2 reveal that mixed language processing temporarily increases phonetic interaction between the languages, causing \nt{2} to converge towards a related L1 category \nt{a:}. The extent and direction of such shift is asymmetrical across the two vowel categories, which parallels findings in prior studies. This asymmetry appears to be mediated by the language-specific distribution of vowel categories. Mixed productions were elicited in two switching paradigms: code-switching and cued picture-naming, to assess whether the postulated discursive elements in code-switching independently influence the outcome of phonetic transfer. Results suggest that in an experimental setup, elicitation through cued-picture-naming vs (controlled) code-switching does not significantly alter the behavior under study.

This study has a number of limitations: (i) the `baseline' values for Bengali vowel categories were estimated from a different set of speakers than the participants of this study, potentially obscuring the role of individual differences in the realization of vowel categories; (ii) we used a frame sentence to elicit code-switched productions such that not all the sentences were pragmatically well-formed. This could have blocked some of the discursive effects that might result from speech planning in ordinary connected speech, thus obscuring differences between the two paradigms under study. Future work should address this by considering data from more naturalistic code-switching paradigms.

@@

\section*{Data availability}
The data files and analysis code are openly available at:  \url{https://osf.io/dsb2x/?view_only=5cb821070bc74a038a293239157c3b11}

@@

\begin{figure}[h] %\centering
	\includegraphics[scale=1]{vowels_e_b_final}
	\caption{Vowel quality in unilinual English (solid lines), mixed condition (dotted lines), and baseline Bengali (colored ellipses) productions}
	\label{vowels_e_b}
\end{figure}

@@

	\includegraphics[scale=1]{vowel_by_subject_ggplot}
	\caption{Subject-wise mean F1; unilingual English}
	\label{boxplot_F1}
\end{figure}

>>>>

Hyderabad,

India\\

@@

Jadavpur University,

Kolkata,

India\\

@@

\newpage
\begin{myparindent}{0pt}
Abstract:\\
This study investigates the phonetic outcome of mixed-language speech in Bengali and Indian English, towards understanding cross-language interaction in highly proficient bilinguals.  We compare spectral properties of L2 vowels \nt{\ae} (common to L1, L2) and \nt{2} (absent in L1) in unilingual vs. mixed productions. Results reveal dynamic and asymmetrical changes in both vowels;  \nt{2} shifting towards a related L1 category \nt{a:}. We interpret this as a temporary increase in cross-language phonetic interaction during mixed-language use, leading to a shift towards L1 norms, evidencing transfer effects on L2 vowels for the first time. We elicit mixed productions through two tasks: cued picture-naming and code-switching, to assess if experimental paradigm independently influences the behavior under study. Results reveal no difference across paradigms. We discuss these findings in light of recent proposals about asymmetries in short-term phonetic interaction, postulated discursive factors in code-switching, and the issue of comparability between paradigms in transfer studies.\\

@@

The rest of this section is structured as follows: \ref{bengali_english_in_india} provides background information on Bengali and Indian English focusing on vowel systems. \ref{causes} discusses two potential sources of transfer during language processing-- global co-activation of multiple languages vs. the local act of switching between languages. Given the highly multilingual setting that characterizes this population, we note that the former is expected to be a constant feature and therefore any observed phonetic interaction is better understood as resulting from the latter. The design of the present study is motivated on the basis of this discussion. A logical consequence of this model is that such phonetic interaction must be highly localized. \ref{duration} elaborates on this by reviewing literature on the duration and `reversibility' of transfer effects. A recurring pattern that emerges from existing research on short-term phonetic transfer is that both languages of the bilingual speaker are not equally affected, although the nature of these reported differences varies greatly across studies. In \ref{asymmetries}, we review some of these findings, focusing on proposed sources of asymmetries, in order to highlight the factors that mediate such phonetic interaction (or fail to do so). We argue that these results point to an urgent need for expanding the range of language pairs and sound categories examined in order to make meaningful generalizations about the nature of short-term phonetic transfer. Following from this, \ref{asymmetry between sounds} discusses reported asymmetries between different sound categories of a single language, and develops the hypotheses of the present study on the basis of these. In \ref{paradigms}, we motivate a secondary aim of this study: to verify the postulated differences between two paradigms for eliciting short-term phonetic interaction in an experimental setup, namely cued picture-naming and code-switching. \ref{questions_and_hypotheses} summarizes the research questions and hypotheses.

@@

Existing research has distinguished changes in category \textsc{representations} due to the acquisition of multiple sound systems (cf. Speech Learning Model \citep{flege1995second,flege2007language}; Perceptual Assimilation Model-L2 \citep{best2007nonnative}), from interaction during accessing, processing, or articulation of these categories (cf. transfer vs. interference \citep{grosjean2012attempt}; competence vs. performance interference \citep{paradis1993linguistic}). 
Given their transient nature, dynamic changes in production during mixed-language use are generally attributed to the latter, e.g. online processing costs \citep{olson2013bilingual,tsui2019impact,vsimavckova2015immediate}, language mode \citep{simonet2014phonetic}, context-awareness \citep{khattab2013phonetic}. What triggers this interaction? \cite{olson2016role} argues that while cross-language phonetic effects are largely measured at the point of language switch, it could have two potential sources:
\begin{enumerate}
	\item The local point of switch itself
	\item Global co-activation of two languages -- \textsc{bilingual language mode} \citep{grosjean1998studying} 
\end{enumerate}

@@

\subsubsection*{L1 vs. L2 status} 
Flege's Speech Learning Model (SLM) \citeyearpar{flege1995second,flege2007language} posits that the sound categories of a bilingual speaker exist in a common phonological space, and therefore in principle, both L1 and L2 categories can influence one another. Thus, the patterns of cross-language transfer depend on how L2 phonemes are mapped in relation to the existing L1 categories. In non-switched production, this is evidenced through a pervasive L1 influence on non-native contrasts that are perceptually linked to an existing L1 contrast (cf. \textsc{equivalence classification}-- \cite{flege1984limits,flege1987production}), and the observation that both L1 and L2 sound systems of bilinguals differ from those of corresponding monolingual speakers \citep[e.g.][]{guion2003vowel}.

@@

The precise contribution of various (linguistic and extra-linguistic) factors and their interactions to the phonetics mixed-language use merits detailed investigation in its own right. In the present study, however, our focus is limited to verifying whether, in an experimental setting, the postulated discursive elements of a CS paradigm alter the extent of observable transfer compared to a paradigm that arguably lacks these--- cued picture-naming. We are less concerned with the exact nature of differences than with their comparability. While researchers have successfully employed each of these paradigms to induce phonetic interaction, comparison across studies is difficult since they also differ in other factors. We provide a starting point for such comparison by eliciting productions of the same tokens from the same set of participants, in both paradigms. A difference in results will suggest that the planning involved in connected speech and code-switching affects transfer above and beyond the phonetic outcome of switching itself, which has implications that future work should consider.

\subsection{Research questions and hypotheses}\label{questions_and_hypotheses}
\begin{itemize}
	\item \textbf{L2 vowels}: Are L2 vowel categories of proficient bilingual speakers affected by dynamic phonetic transfer during mixed-language production, as evidenced by differences in the spectral properties of L2 vowels in switched and non-switched tokens?
	\item \textbf{Asymmetry}: If so, we expect an asymmetry between the vowels in the extent and direction of shift. Based on comparisons with Bengali vowels from \cite{dutta2021} we expect \nt{2} to be lowered and fronted in the mixed condition. \nt{\ae} is expected to show lowering, but to a lesser extent than \nt{2}, as it is in a part of the IE vowel space that has greater vowel density.
	\item \textbf{Paradigm}: Is there an observable difference between the extent of transfer in a cued picture-naming vs. a code-switching paradigm? 
\end{itemize}

@@

10 Bengali-English bilingual speakers (male=5, age range 19 to 28), completed the tasks. All the participants speak other languages in addition to Bengali and English. At the time of recording, they were living in or around the EFL University campus in Hyderabad, which has a linguistically diverse population. Speakers were selected on the basis of responses to a Language Background Questionnaire to ensure comparable LSRW (listening, speaking, reading, writing) skills in L1 and L2 across participants, and were compensated for their time.

\subsection{Stimuli}
Twenty monosyllabic English words containing the vowels \nt{2} or \nt{\ae} were selected as target words. To minimize coarticulatory effects on the vowel, the pre-vocalic consonants were all voiced plosives (\nt{b} or \nt{d}). 10 monosyllabic English words that do not contain the target vowel served as filler items. We avoided English words that are lexicalized loanwords in Bengali. %The list of words can be found in sec.\ref{appendix}.

@@

\begin{enumerate}[]
	\item Cued picture-naming: Participants were presented with slides in the following sequence:
	\begin{itemize}
		\item Language cue- a word in either English or Bengali orthography
		\item A picture displayed for 50ms; named by the participant in either English or Bengali, depending on the language cue 
		\item Target -- an English word; read out by the participant
		\item Distracter math problem
	\end{itemize}
	Since all the target words were in English, the language cue was used to manipulate the \textsc{preceding} word, giving target words in non-switched (unilingual) and switched (mixed) utterances. The orthographies of English and Bengali are visually distinct, and were therefore used to cue language.

@@

\end{enumerate}
\subsection{Acoustic analysis}\label{analysis}
The first iteration of each task was treated as a trial, and excluded from further analyses. An additional 67 tokens (2.79\%) were excluded due to errors in presentation/production, giving 2333 target word tokens for analysis. Recordings were manually segmented and annotated in Praat \citep{boersma2016praat}, and the first and second formant frequencies (F1 and F2) of the target vowels \nt{2, \ae} were measured at five points-- in 10\% increments starting at 5\% into the vowel. This gave measures of F1 and F2 at 5\%, 15\%, 25\%, 35\% and 45\% into the vowel, allowing us to observe any dynamic changes in the extent of transfer over the course of producing the vowel.

@@

Vowel quality and gender are expected to systematically affect F1. Thus, we began with a model that had Vowel and Gender as fixed effects, random intercepts for Subject and Item, and random slopes for each within-unit variable that was relevant to our hypotheses: by-subject slopes for Task, Context, and Vowel; by-item slopes for Task and Context. Next, adding Context as a fixed effect significantly improved the model fit ($\chi^2(1) = 5.54, p= .01854$), suggesting that the experimental condition affects vowel height. Following from our hypothesis \textbf{Asymmetry}, we expected the extent of change in vowel height between the conditions to be mediated by the identity of the vowel. Thus, we added a two-way interaction factor between Context and Vowel (Context*Vowel). As expected, this was found to significantly improve model fit ($\chi^2(1) = 12.61, p= .0003837$). To assess whether the experimental paradigm affected F1, we added Task as a fixed effect. This marginally improved the model ($\chi^2(1) = 3.74, p= .05302$). Towards addressing our research question \textbf{Paradigm}, we wanted to assess whether the extent of shift differs across the two paradigms. However, no significant improvement was achieved by adding an interaction effect between Task and Context. Subsequent additions of other interactions, and removing each of the fixed effects in turn, did not lead to any further improvement.

@@

With regard to our first research question \textbf{L2 vowels}, the results confirm that the L2 vowels of proficient bilinguals can be affected by mixed-language processing. Both vowels in this study showed systematic shifts in the mixed-language condition compared to baseline unilingual productions. In the context of an ICM--based analysis of between-language asymmetries in phonetic transfer as proposed by \cite{olson2013bilingual}, this is a significant finding: it demonstrates that the postulated lower switch cost on an L2 does not entirely mask transfer effects on L2 vowels. Moreover, our findings differ from \cite{tsui2019impact} who report on the basis of VOT measurements that balanced bilinguals do not evidence an effect of language-switching. \citeauthor{tsui2019impact} attribute the absence of transfer effects to greater experience with language switching in these participants, leading to lower switch-costs in both languages. We find that even proficient bilinguals with extensive experience of language switching demonstrate cross-language transfer effects in vowel production. Thus, following from the discussion in sec. \ref{asymmetries}, it might be fruitful for future studies to focus on vowels as a potential site for transfer. This would both expand the empirical scope of existing research (cf. sec. \ref{asymmetry between sounds}) and avoid confounding factors arising from independent differences between long- and short-lag VOT languages. More generally, these findings point to a need for expanding the range of features in which transfer effects are studied, in order to make robust generalizations about the mechanisms underlying observed patterns.

@@

Our third research question \textbf{Paradigm} concerned the effect of experimental paradigm on the phonetic outcome of mixed-language use. \cite{olson2013bilingual} proposes that in both scripted and spontaneous code-switching paradigms, ``discursive properties may partially drive the production of code-switches, masking underlying effects of interaction", and thus proposes cued picture-naming as a preferable alternative. If such discursive properties independently affect the phonetic outcome of language mixing in an experimental setup, this calls to question the comparability of results across studies using these different paradigms. We verified this by comparing productions of the same items from the same group of participants in a scripted CS paradigm and a cued picture-naming paradigm. We found no significant interaction of Task either with Context, or with the Vowel*Context interaction, suggesting that the extent of phonetic transfer was not affected by the paradigm. 
There are at least two plausible reasons for this-- (i) it is possible that the planning effects of CS do not translate to an increase/decrease in observable transfer at all. This would mean that the pre-switch effects observed, for example, in \cite{muldner2019phonetics} and \cite{bullock2009trying} are better interpreted as resulting from some other aspect of the experimental setup, than the act of code-switching itself. While planning could additionally affect the words preceding the switch, the effects on the switched token itself do not change. (ii) Our study used a frame sentence to induce code switching, and not all the items led to pragmatically meaningful sentences. Thus, another possible explanation is that the discursive effects discussed by \cite{olson2013bilingual} are triggered by more naturalistic paradigms where attention to meaning and interlocutor are important. More within-study comparisons with less structured paradigms, and spontaneous code-switching data, can clarify this. The results of the present study indicate that for the purposes of observing dynamic phonetic transfer in an experimental setup, elicitation through cued-picture-naming vs. (controlled) code-switching does not significantly alter the behavior under study. Thus, the choice of paradigm can be based on other considerations. Moreover, this suggests that (all else being equal) the results from such studies can be compared directly.

@@

\section{Conclusion}
This study examined the phonetic outcome of mixed-language production in a group of highly proficient bilingual speakers of Bengali and Indian English. Specifically, we examined the spectral quality of two L2 vowel categories \nt{2} and \nt{\ae} produced in two experimental conditions: unilingual (English) and mixed (switching from Bengali to English).
Comparisons of F1 and F2 reveal that mixed language processing temporarily increases phonetic interaction between the languages, causing \nt{2} to converge towards a related L1 category \nt{a:}. The extent and direction of shift is asymmetrical across the two vowel categories, which parallels findings in prior studies. This asymmetry appears to be mediated by the language-specific distribution of vowel categories. These findings evidence an effect of phonetic transfer on L2 vowels for the first time, and raise questions for proposed inhibition-based accounts of cross-language phonetic interaction. Mixed productions were elicited in two switching paradigms: code-switching and cued picture-naming, to verify whether the postulated discursive elements in code-switching independently influence the outcome of phonetic transfer. Results suggest that in an experimental setup, elicitation through cued-picture-naming vs. (controlled) code-switching does not significantly alter the behavior under study.

This study has a number of limitations: (i) the `baseline' values for Bengali vowel categories were estimated from a different set of speakers than the participants of this study, potentially obscuring the role of individual differences in the realization of vowel categories; within-speaker comparisons of L1 and L2 vowels is likely to offer a more fine-grained view of the direction of shift in relation to L1 categories.  (ii) we used a frame sentence to elicit code-switched productions such that not all the sentences were pragmatically well-formed. This could have blocked some of the discursive effects that might result from speech planning in ordinary connected speech, thus obscuring differences between the two paradigms under study. Future work should address this by considering data from more naturalistic code-switching paradigms.

@@

\section*{Data availability}
The materials, data files, and analysis code for this study are openly available in OSF at:  \url{https://doi.org/10.17605/OSF.IO/DSB2X}.

@@

\begin{figure}[h] %\centering
	\includegraphics[scale=1]{vowels_e_b_final}
	\caption{Vowel quality in unilinual English, mixed, and baseline Bengali productions}
	\label{vowels_e_b}
\end{figure}

@@

	\includegraphics[scale=1]{vowel_by_subject_ggplot}
	\caption{Subject-wise mean F1 in unilingual English productions}
	\label{boxplot_F1}
\end{figure}
